---
title: "ForceAtlas2 layout for network analysis"
author: "Adolfo Ãlvarez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ForceAtlas2}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This is the R implementation of the Force Atlas 2 graph layout designed for Gephi. The algorithm is detailed in:

[Jacomy M, Venturini T, Heymann S, Bastian M (2014) ForceAtlas2, a Continuous Graph Layout Algorithm for Handy Network Visualization Designed for the Gephi Software. PLoS ONE 9(6): e98679](http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0098679)


## Installation

This package is not yet available in CRAN, so install it directly from Github with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("analyxcompany/ForceAtlas2")
```

## Usage

After installation the package is loaded as usual with:

```{r}
library(ForceAtlas2)
```

This implementation accepts as inputs an `igraph` object or a data frame. A full parameter description of the algorithm can be found in [Jacomy et al](http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0098679), but a summary is accesible via `help("layout.forceatlas2")`.

Among those parameters you would probably like to set the number of iterations (`iterations`) and how often a plot of the positions should be generated (`plotstep`). This is particularly useful to evaluate the convergence of the algorithm. Set `plotstep=0` to suppress intermediate plots.

**igraph input**

For this example I will use the coappeareance network from Les Miserables, by Victor Hugo. Get more details about this data set with `igraph::nexus.info("miserables")`

```{r, echo=TRUE, eval=FALSE, fig.width = 7.2, fig.height = 6}
library(igraph) 
g <- nexus.get("miserables")
layout <- layout.forceatlas2(g, iterations=2000, plotstep=100)
```

```{r, echo=FALSE, fig.width = 7.2, fig.height = 6}
library(igraph) 
g <- nexus.get("miserables")
layout <- layout.forceatlas2(g, iterations=2000, plotstep=2000)
```

This can be also plotted from as an igraph object:

```{r, fig.width = 7.2, fig.height = 6}
plot(g, layout=layout)
```

**data frame input**

For those not familiar with the `igraph` package, is possible to calculate the algorithm directly from a data frame. This data frame should consist in three columns: `from`, `to`, and `weights`, indicating the corresponding nodes connections and the weights. 

```{r}
data <- get.data.frame(g)
dim(data) #The dimension of the data
head(data,10) #We show just the first 10 rows
```

After you have your data in that format, the application of the function is equivalent to the previous one, with one exception, the parameter `directed` indicating if the network is directed or not (`directed = TRUE` by default). In this example, `the miserables` is an undirected graph so we change it accordingly.

```{r, eval=FALSE, fig.width = 7.2, fig.height = 6}
layout.forceatlas2(data, directed=FALSE, iterations = 2000, plotstep = 100)
```

```{r, echo=FALSE, fig.width = 7.2, fig.height = 6}
layout <- layout.forceatlas2(data, directed=FALSE, iterations = 2000, plotstep = 2000)
```

When the input is a data frame, the output is also a data frame including the name of the nodes and positions.

```{r}
head(layout, 10) #Show just the 10 first rows
```
